Mean temperature, fluctuation magnitude, and level of biological organization affect biological responses amidst thermal variability: a meta-analysis

### 2021-05-11 writing /thesis_draft/03-chap3.Rmd
putting in everyone's edits from orals

```{r, echo= FALSE, message= FALSE, fig.cap="Forest plot of effects sizes by study and experiment within a study. Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SMD for multiple responses within each experiment in a single study. Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.", fig.scap="Forest plot of effect sizes", out.width="95%", fig.align="center"}
forest <- dat_MA_ES %>%
  group_by(study_id, experiment_id) %>%
  summarise(mean_yi = mean(yi),
            mean_vi = mean(vi)) %>%
  unite(group, c(study_id, experiment_id), sep = " experiment ", remove = FALSE)
forest(forest$mean_yi, forest$mean_vi, slab=forest$group, pch =19, psize=1, col = "gray2", 
       header=TRUE, cex=0.77)
```

Overall, even our simple model (n=140) demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p < 0.0001), as demonstrated by Figure 3.1. These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis (see Appendix Figure 1). As was referenced in the methods, this allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.

Our full model, including 7 covariates (all, an additional interaction term between fluctuation range and mean temperature, and excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM = 191.3722, df = 7, p < 0.0001). Further, fluctuation range (n=140) and the interaction term between fluctuation range and mean temperature (n=140) were statistically significant in the model (df = 7, p < 0.00001). Organization level (n=140) was also important in the model, as a statistically significant covariate (df= 7, p < 0.05).

```{r,  echo=FALSE, results = 'hide', message=FALSE, fig.cap = "Coefficient estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.  Error bars represent 95 percent confidence intervals.", out.width="95%", fig.align="center", fig.scap="Coefficient estimates for full model"}
full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level + resp_type, 
               random = ~1 |  study_id/ experiment_id/ response_id,
                 method="REML")

@@

Fluctuation range (n=140) had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p < 0.001 level (Figure 3.2). The most negative estimate, organization level (n=140), was significantly influential in our model at a significance level of p < 0.05. Surprisingly, the additional negative estimates of our model coefficients, experimental age (n=140) and size (n=140), were not statistically significant. In our analysis of thermal stress via a separate random effects model, we found thermal stress (n=132) was not statistically significant (df=1, p =0.4855) in explaining the effect size (Table 3.1).

@@

The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 3.3. At lower mean temperatures, higher fluctuation ranges have a more positive effect on organism responses and performance, though that trend starkly ends at about $24^{\circ}$C. Beyond $24^{\circ}$C, the effect size decreases to nearly half the size of colder temperatures and hovers just above an effect size of one (Figure 3.3).

```{r, echo=FALSE, message = FALSE, fig.cap="Scatterplot describing the relationship between mean temperature, temperature fluctuation range, and effect size. Temperature fluctuation ranged from 0-20°C, Mean temperature ranged from 7-33°C. Effect size ranged from -4 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure 2).", fig.align="center", out.width = '90%', fig.scap="Scatterplot of relationship between range, mean, and effect size"}
normalized %>%
ggplot(aes(x=mean_temp_constant, y = yi, fill=flux_range))+
  geom_jitter( alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
  scale_fill_viridis('Temperature range (°C)', option = "C")+
  theme_classic()+
  xlab("Mean Temperature (°C)")+
  ylab("Effect Size")+
  geom_line(y=0, linetype = "dashed")+
  theme(legend.position="bottom")
```

@@

As noted in the methods, when we excluded one highly influential study and included thermal stress in the exclusion model (n=132), size became much more important in the exclusion model (df=8, p < 0.001) as did mean temperature (df=8, p < 0.001). Surprisingly, organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the exclusion model (df=8, p < 0.1).

@@

The results of this meta-analysis suggest that fluctuation range is of significant importance to performance, both at the individual and population level. There were not overall differences in the direction of thermally variable environments positively or negatively impacting performance relative to constant environments. However, a considerable amount of variation exists in whether thermally variably environments positively, negatively, or minimally impact performance across studies. Our model results highlight the importance of fluctuation range, mean temperature, and organization level to better understand the variety of effects thermal variation has on organismal responses. These results suggest that mean temperature and fluctuation range must be considered together when understanding the impacts of environmental variability, i.e. higher fluctuation ranges at lower mean temperatures have higher effect sizes compared to effects sizes at lower fluctuation ranges at high mean temperatures. The ways in which organisms, populations, and communities have dealt with natural disturbances has become even more significant in the face of climate change [@vasseur_increased_2014; @sunday_thermal_2012]. With global mean temperatures expected to exceed the crucial 2 degree centigrade threshold scientists have deemed “the point of no return” [@russill_tipping_2009], understanding how organisms will cope is key for understanding how ecosystem composition will change at all scales in the next several decades [@west_allometric_2002; @cheung_application_2008]. In this context, perhaps the range of temperatures experienced is of greater importance to predictions of performance under future conditions than simply mean temperature [@vasseur_increased_2014].

These findings underscore the significance of better understanding why differences across studies and experimental designs occurred. Jensen’s inequality is one way to conceptualize why responses differ, as averaging nonlinear responses linearly will not accurately predict anticipated performance, in fact, it will underestimate performance [@ruel_jensens_1999; @bernhardt_nonlinear_2018]. The ramifications of underestimating performance and not accounting for environmental variability are inumerable, the biggest of which may be our inability to gauge the critical tipping point for many species' ability to survive. Acclimation may also explain differences in response across studies as the scale on which variation occurs influences how organisms may respond. If organisms’ non-genetic phenotypic changes lag behind environmental changes but outpace changes in the constant environment (i.e. inducing gradual plasticity), acclimation duration may explain mismatches in responses across studies [@kremer_gradual_2018]. In terms of variability pattern, if an organism experiences environmental variability in uneven intervals, acclimation to previous conditions could explain increases or decreases in performance. Rapid evolution could further explain differences in responses across studies. If organisms are able to keep evolutionary pace with environmental changes via genetic phenotypic changes, responses could be positively affected, as demonstrated in predator-prey system dynamics [@yoshida_rapid_2003]. It is important to note that rapid evolution will be particularly relevant to population levels responses and not individual level responses.

@@

An important side note is that our full model and thermal stress model differed in the number of studies, as one study, @garcia-ruiz_effects_2011, did not have any supplemental information on thermal performance curve points for the taxa used in that study (Xylotrechus arvicola). Further, this study happened to contribute highly influential points to our model and subsequent analysis, as our model results change drastically when excluding @garcia-ruiz_effects_2011 from our full model. We decided to include this study in our analysis because data were extracted from a table, therefore there were no inaccuracies in obtaining the data. It provides an important counter and discussion for how and why responses from different studies may be so different from others.

@@

We also expected for experimental age to be an important predictor, as the time period in which an organism experiences thermal variability has been demonstrated to be important for organisms like turtles, relying on TSD for development (Bowden and Paitz 2018). This may again be the result of many studies not focused on explicitly the larval stage of development, mainly focusing on the juvenile stage instead. However, there have been meta-analyses, both qualitative [@massey_measurement_2019], and quantitative [@noble_developmental_2018], that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response [@noble_developmental_2018]. These results align with patterns of juvenile organisms (i.e. fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance [@portner_ecology_2008; @moyano_effects_2017].

Beyond the data included in our meta-analysis, it is also important to note that there was little variety in the pattern of variation (diurnal, colored noise, etc.) as was initially a question of interest for this project. As Figure 1.3 demonstrated, diurnal cycles are more correlated than reddened cycles, like seasonal temperatures. It is surprising that so many thermal variability studies are focused on diurnal cycles, when in fact, temperature patterns are more stable than diurnal cycles, and less autocorrelated. Are these diurnal temperature fluctuations accurate for what is expected under natural conditions? This lack of variety in pattern may be due to a lack of consensus on the language used to classify and discuss thermal variability. Though there is a large body of literature that investigates temperature variation, there appears to be dissonance in the scientific literature about the language used to discuss such variation. Many of the background papers pulled from the systematic literature search used temperature variation to mean just different temperatures [@amarasekare_framework_2012] as well as different ranges of temperature, both of which are different concepts. For instance, @amarasekare_framework_2012 addressed the need for more complexity when we think about the processes underlying temperature dependent fitness, as fitness is an amalgamation of interactions between different life-history traits. Though temperature variation appeared in the keywords of the article, there was little to no mention of actual thermal variation or variation pattern. Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly irrelevant thermal regimes [@colinet_mechanisms_2018; @colinet_insects_2015]. Using standardized language to describe specific patterns or durations of variability may be helpful in better summarizing the results and conclusions from previous studies.

@@

Several studies made little to no mention of generation time with respect to duration of their experiments or study organisms used. How organisms’ life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond [@bernhardt_life_2020]. This is also pertinent to scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta-analysis. Diurnal thermal fluctuations may affect organisms with shorter generation times differentially compared to organisms with longer generation times. The converse is true for seasonal or decadal thermal fluctuations.

>>>>

```{r, echo= FALSE, message= FALSE, fig.cap="Forest plot of effects sizes by study and experiment within a study. Data indicate that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SMD for multiple responses within each experiment in a single study. Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.", fig.scap="Forest plot of effect sizes", out.width="95%", fig.align="center"}
forest <- dat_MA_ES %>%
  group_by(study_id, experiment_id) %>%
  summarise(mean_yi = mean(yi),
            mean_vi = mean(vi)) %>%
  unite(group, c(study_id, experiment_id), sep = " experiment ", remove = FALSE)
forest(forest$mean_yi, forest$mean_vi, slab=forest$group, pch =19, psize=1, col = "gray2", 
       header=TRUE, cex=0.77)
```

Overall, even our simple model demonstrated that the intercept was not significant but there was significant heterogeneity across studies (n=140, z= 0.79, p < 0.0001), as demonstrated by Figure 3.1. These results allowed us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis (see Appendix Figure 1). This then allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.

Our full model, including 7 covariates (all covariates in Table 2.2, an additional interaction term between fluctuation range and mean temperature, and excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM = 191.3722, df = 7, p < 0.0001). Further, fluctuation range  and the interaction term between fluctuation range and mean temperature  were statistically significant in the model (n=140, df = 7, p < 0.00001). Organization level was also important in the model, as a statistically significant covariate (n=140, df= 7, p < 0.05).

```{r,  echo=FALSE, results = 'hide', message=FALSE, fig.cap = "Coefficient estimates for each covariate in the full model, where * signify statistically significant model predictors.  Error bars represent 95 percent confidence intervals. Yellow indicates a positive coefficient estimate, purple indicates a negative coefficient estimate.", out.width="95%", fig.align="center", fig.scap="Coefficient estimates for full model"}
full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level + resp_type, 
               random = ~1 |  study_id/ experiment_id/ response_id,
                 method="REML")

@@

Fluctuation range had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant (n=140, df=7, p < 0.001 level (Figure 3.2). The most negative estimate, organization level (n=140), was also significantly influential in our model (n=140, df=7, p < 0.05).The additional negative estimates of our model coefficients, experimental age and size, were not statistically significant. In our analysis of thermal stress via a separate random effects model, we found thermal stress was not statistically significant (n=132, df=1, p = 0.4855) in
explaining the effect size (Table 3.1).

@@

The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 3.3. At lower mean temperatures, higher fluctuation ranges generally have a more positive effect on organism responses and performance, though that trend starkly ends at about $24^{\circ}$C. However, at higher mean temperatures, lower fluctuation ranges appears to have an equally positive and negative effect on organism responses and performance (Figure 3.3).

```{r, echo=FALSE, message = FALSE, fig.cap="Scatterplot describing the relationship between mean temperature, temperature fluctuation range, and SMD. Temperature fluctuation ranged from 0-20°C, Mean temperature ranged from 7-33°C. SMD ranged from -4 to 6, restricted based on the distribution of SMD to minimize impact of outliers (see Appendix Figure 2).", fig.align="center", out.width = '90%', fig.scap="Scatterplot of relationship between range, mean, and SMD"}
normalized %>%
ggplot(aes(x=mean_temp_constant, y = yi, fill=flux_range))+
  geom_jitter(alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
  scale_fill_viridis('Temperature range (°C)', option = "C")+
  theme_classic()+
  xlab("Mean Temperature (°C)")+
  ylab("SMD")+
  geom_line(y=0, linetype = "dashed")+
  theme(legend.position="bottom")
```

@@

When we excluded one highly influential study and included thermal stress in the exclusion model (n=132), size became much more important in the exclusion model (df=8, p < 0.001) as did mean temperature (df=8, p < 0.001). Organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the exclusion model (df=8, p < 0.1).

@@

The results of this meta-analysis suggest that fluctuation range is of significant importance to performance, both at the individual and population level. There were not overall differences in the direction of thermally variable environments positively or negatively impacting performance relative to constant environments. However, a considerable amount of variation exists in whether thermally variably environments positively, negatively, or minimally impact performance across studies. Our model results highlight the importance of fluctuation range, mean temperature, and organization level to better understand the variety of effects thermal variation has on organismal responses. These results suggest that mean temperature and fluctuation range must be considered together when understanding the impacts of environmental variability, i.e. higher fluctuation ranges at lower mean temperatures generally have higher effect sizes compared to higher and lower effects sizes of lower fluctuation ranges at higher mean temperatures. The ways in which organisms, populations, and communities have dealt with natural disturbances has become even more significant in the face of climate change [@vasseur_increased_2014; @sunday_thermal_2012]. With global mean temperatures expected to exceed the crucial 2 degree centigrade threshold scientists have deemed “the point of no return” [@russill_tipping_2009], understanding how organisms will cope is key for understanding how ecosystem composition will change at all scales in the next several decades [@west_allometric_2002; @cheung_application_2008]. In this context, perhaps the range of temperatures experienced is of greater importance to predictions of performance under future conditions than simply mean temperature [@vasseur_increased_2014].

These findings underscore the significance of better understanding why differences across studies and experimental designs occurred. Jensen’s inequality is one way to conceptualize why responses differ, as averaging nonlinear responses linearly will not accurately predict anticipated performance, in fact, it will underestimate performance [@ruel_jensens_1999; @bernhardt_nonlinear_2018]. The ramifications of underestimating performance and not accounting for environmental variability are inumerable, the biggest of which may be our inability to gauge the critical tipping point for many species' ability to survive. Acclimation may also explain differences in response across studies as the scale on which variation occurs influences how organisms may respond. If organisms’ non-genetic phenotypic changes lag behind environmental changes but outpace changes in the constant environment (e.g. inducing gradual plasticity), acclimation duration may explain mismatches in responses across studies [@kremer_gradual_2018]. In terms of variability pattern, if an organism experiences environmental variability in uneven intervals, acclimation to previous conditions could explain increases or decreases in performance. Rapid evolution could further explain differences in responses across studies. If organisms are able to keep evolutionary pace with environmental changes via genetic phenotypic changes, responses could be positively affected, as demonstrated in predator-prey system dynamics [@yoshida_rapid_2003]. It is important to note that rapid evolution will only be relevant to population levels responses and not individual level responses.

@@

An important side note is that our full model and thermal stress model differed in the number of studies, as one study, @garcia-ruiz_effects_2011, did not have any supplemental information on thermal performance curve points for the taxa used in that study (*Xylotrechus arvicola*). Further, this study happened to contribute highly influential points to our model and subsequent analysis, as our model results change drastically when excluding @garcia-ruiz_effects_2011 from our full model. We decided to include this study in our analysis because data were extracted from a table, therefore there were no inaccuracies in obtaining the data. It provides an important counter and discussion for how and why responses from different studies may be so different from others.

@@

We also expected for experimental age to be an important predictor, as the time period in which an organism experiences thermal variability has been demonstrated to be important for organisms like turtles, relying on TSD for development (Bowden and Paitz 2018). This may again be the result of studies not focused on the larval stage of development, but instead on the juvenile stage. However, there have been meta-analyses, both qualitative [@massey_measurement_2019], and quantitative [@noble_developmental_2018], that explicitly looked at the effects of incubation temperature on reptiles, concluding that there is a moderate to large effect of incubation temperature on the magnitude of response [@noble_developmental_2018]. These results align with patterns of juvenile organisms (e.g. fish) routinely having higher CTmax values than adults, such that life stage during experimental duration is of importance to thermal tolerance and performance [@portner_ecology_2008; @moyano_effects_2017].

Beyond the data included in our meta-analysis, it is also important to note that there was little variety in the pattern of variation (diurnal, colored noise, etc.) as was initially a question of interest for this project. Diurnal cycles are more correlated than reddened cycles, like seasonal temperatures (Figure 1.3). It is surprising that so many thermal variability studies are focused on diurnal cycles, when in fact, longer-term temperature patterns are more stable than diurnal cycles, and less autocorrelated. Are these diurnal temperature fluctuations accurate for what is expected under natural conditions? This lack of variety in pattern may be due to a lack of consensus on the language used to classify and discuss thermal variability. Though there is a large body of literature that investigates temperature variation, there appears to be dissonance in the scientific literature about the language used to discuss such variation. Many of the background papers pulled from the systematic literature search used temperature variation to mean just different temperatures [@amarasekare_framework_2012] as well as different ranges of temperature, both of which are different concepts. For instance, @amarasekare_framework_2012 addressed the need for more complexity when we think about the processes underlying temperature dependent fitness, as fitness is an amalgamation of interactions between different life-history traits. Though temperature variation appeared in the keywords of the article, there was little to no mention of actual thermal variation or variation pattern. Additionally, a subsection of the papers included in our search specifically defined thermal variability as short term changes in temperature, though these studies were mainly focused on explicitly ecologically irrelevant thermal regimes [@colinet_mechanisms_2018; @colinet_insects_2015]. Using standardized language to describe specific patterns or durations of variability may be helpful in better summarizing the results and conclusions from previous studies.

@@

Several studies made little to no mention of generation time with respect to duration of their experiments or study organisms used. How organisms’ life cycles and life spans interact with and are impacted by fluctuation patterns is important for understanding the layers variation permeates and how organisms respond [@bernhardt_life_2020]. This is also pertinent to the scale at which we study variability, given the overrepresentation of diurnal fluctuation patterns featured in our meta-analysis. Diurnal thermal fluctuations may affect organisms with shorter generation times differentially compared to organisms with longer generation times. The converse is true for seasonal or decadal thermal fluctuations.

### 2021-05-11 writing /thesis_draft/05-appendix.Rmd
putting in everyone's edits from orals

```{r, echo= FALSE}
table4 <- data.frame(Species = c("Drosophila melanogaster", "Leptopilina boulardi", 
                         "Mauremys reevesii", "Gryllus firmus", 
                         "Limnodynastes peronii", "Limnodynastes tasmaniensis", 
                         "Platyplectrum ornatum ", "Natrix natrix", 
                         "Coturnix japoncia", "Plestiodon chinensis", 
                         "Rhodnis prolixus", "Escherichia coli", 
                         "Salmonella", "Caiman latirostris", 
                         "Trachemys scripta", "Esteya vermicola"), 
                 Topt = c("24-26.6 C", "25 C", "37.97 +/- 0.64", "< 30 C", "30 C", "30-33 C",
                          "33 C", " </= 30 C", "26 C", "33 C", "30-35 C", "37 C, 35-36 C",
                          "35-36 C", "32-33 C", "32-33 C", "26 C"), 
                 `Reference` = c("Klepsatel 2013, David 1988", "De lava et al 2016", 
                                 "Dang et al 2019", "Singh et al 2020", "Seebacher et al 2014",
                                 "Whitehead et al 1989", "Kern et al 2014", "Issac 1997",
                                 "Alagawany et al 2017", "Baojun et al 2010", "Fresquet and Lazzari 2011",
                                 "Doyle and Schoeni 1984, Bronikowski et al 2001", 
                                 "Bronikowski et al 2001", 
                                 "Simoncini et al 2019, Parachu-Marco et al 2017, Pina et al 2002", 
                                 "Dang et al 2019", "Wang et al 2015"))
kable(table4, "latex", booktabs = T, 
      col.names = c("Species", "Topt", "Reference"), escape = F, caption = "TPC metrics extracted from additional papers to collect thermal history data for each species in the meta analysis (missing data on species X. arvicola, used in Garcia-Ruiz et al 2011)", caption.short ="Thermal stress data by species") %>%
   row_spec(0,bold=TRUE) %>%
  column_spec(3, width = "6cm") %>%
   kable_styling(latex_options = c("striped", "hold_position"))
```

>>>>

```{r, echo= FALSE}
table4 <- data.frame(Species = c("Drosophila melanogaster", "Leptopilina boulardi", 
                         "Mauremys reevesii", "Gryllus firmus", 
                         "Limnodynastes peronii", "Limnodynastes tasmaniensis", 
                         "Platyplectrum ornatum ", "Natrix natrix", 
                         "Coturnix japoncia", "Plestiodon chinensis", 
                         "Rhodnis prolixus", "Escherichia coli", 
                         "Salmonella", "Caiman latirostris", 
                         "Trachemys scripta", "Esteya vermicola", "Xylotrechus arvicola"), 
                 Topt = c("24-26.6°C", "25°C", "37.97 ± 0.64 °C", "30°C", "30°C", "30-33°C",
                          "33°C", "30°C", "26°C", "33°C", "30-35°C", "37°C, 35-36°C",
                          "35-36°C", "32-33°C", "32-33°C", "26°C", "N/A"), 
                 Class = c("Insecta", "Insecta", "Reptilia", "Insecta", "Amphibia", "Amphibia", "Amphibia",
                           "Reptilia", "Aves", "Reptilia", "Insecta", "Gammaproteobacteria",
                           "Gammaproteobacteria","Reptilia", "Reptilia", "Not assigned; Fungus", "Insecta"),
                 `Reference` = c("Klepsatel 2013, David 1988", "De lava et al 2016", 
                                 "Dang et al 2019", "Singh et al 2020", "Seebacher et al 2014",
                                 "Whitehead et al 1989", "Kern et al 2014", "Issac 1997",
                                 "Alagawany et al 2017", "Baojun et al 2010", "Fresquet and Lazzari 2011",
                                 "Doyle and Schoeni 1984, Bronikowski et al 2001", 
                                 "Bronikowski et al 2001", 
                                 "Simoncini et al 2019, Parachu-Marco et al 2017, Pina et al 2002", 
                                 "Dang et al 2019", "Wang et al 2015", "Garcia-Ruiz et al 2011"))
kable(table4, "latex", booktabs = T, 
      col.names = c("Species", "Topt (°C)", "Class", "Reference"), escape = F, caption = "TPC metrics extracted from additional papers to collect thermal history data for each species in the meta-analysis (missing data on species X. arvicola, used in Garcia-Ruiz et al 2011) and scientific classification at the class level", caption.short ="Thermal stress data by species") %>%
   row_spec(0,bold=TRUE) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(4, width = "3cm") %>%
   kable_styling(latex_options = c("striped", "hold_position"))
```
